---
title: "Monty Hall Monte Carlo"
author: "Rachel Meade"
date: "8/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Monty Hall and Monte Carlo
Whether you recognize its name or not, you've probably heard of the Monty Hall problem. It's a probability puzzle, loosely based on the American television game show "Let's Make a Deal" and named after its original host, Monty Hall.

Suppose you're on a game show, and you're given the choice of three doors: Behind one door is a car; behind the others, goats. You pick a door, say No. 1, and the host, who knows what's behind the doors, opens another door, say No. 3, which has a goat. He then says to you, "Do you want to pick door No. 2?" Is it to your advantage to switch your choice? 

Instead of evaluating this question by using probability calculus, let's try exploring using Monte Carlo simulations. 

```{r Small Monte Carlo}
N = 10 # Suppose you play the game 10x
prize = 1 # The prize is behind the door #1.
doors = c(1,2,3) # There are three doors #1, #2, and #3.
first = sample(doors, size = N, replace=TRUE) # Suppose your pick is purely random.
first
host = 1+2+3 - prize - first # If you pick the wrong door, the host will open the only remaining door.
host # Of course, sometimes you pick the right door.
N.right = sum(first == prize) # This is the number of times you pick the right door at first.
N.right
remain.doors = setdiff(c(1,2,3),prize) # This is a set of doors without the prize.
remain.doors
# If you chose the right door, the host will randomly choose a door between two remaining doors #2 and #3.
host[first == prize] =  sample(remain.doors, size = N.right, replace=TRUE) 
cbind(prize,first,host) # See what door the host opens.
second = 1+2+3-first-host # If you switch, you pick the only remaining door.
cbind(prize,first,host,second) # This is the results of 10 games.
mean(prize==first)
mean(prize==second)
```

Now let's repeat this game 1 million times instead of just 10.

```{r}
N = 10^6  
prize = 1
doors = c(1,2,3)
first = sample(doors, size = N, replace=TRUE) 
host = 1+2+3 - prize - first
N.right = sum(first == prize)
remain.doors = setdiff(c(1,2,3),prize)
host[first == prize] =  sample(remain.doors, size = N.right, replace=TRUE)
second = 1+2+3-first-host
mean(prize==first)
mean(prize==second)
```

Even if you change where the prize is (that is, prize = 2 or 3), the answer is the same. In fact, it shouldn't matter because we could just rename the door. 
 
Still skeptical? You can also find the right answer by logic. 

Suppose you play this game 1 million times.
- If you don't switch, the probability of winning is 1/3.
- Now suppose there is a guy who plays next to you every time. 
= He follows your first choice and then he switches always when he is asked by the host.
- Then, at each game, either you or he will get the prize. 
- If you get the prize one of three games, it means he gets the prize two out of three games.

Therefore, the player who switches always has 2/3 winning probability while the one who keeps the first choice always has 1/3 winning probability.

```

